---
title: "Getting started with scAgentKit on PBMC 3k"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with scAgentKit on PBMC 3k}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE   # 演示用,CI 不实际跑;读者本地有 SeuratData 时可手动 eval
)
```

## What this vignette covers

This walkthrough runs scAgentKit end-to-end on the standard PBMC 3k
dataset (~2,700 cells, 13 known immune populations). The point is to
show the *shape* of the pipeline — what every step looks like, what
gets recorded, and how to inspect the decision log and reproducible
script when the analysis finishes.

Because the LLM calls cost real money and require an API key, the
chunks below are not evaluated when the vignette is built. Copy them
into an R session with `ANTHROPIC_API_KEY` (or any other supported
provider's key) set in `~/.Renviron`.

## 1. Install and load

```{r install}
# install.packages("remotes")
# remotes::install_github("ChanghaoKan/scAgentKit")
# remotes::install_github("satijalab/seurat-data")

library(scAgentKit)
library(Seurat)
library(SeuratData)
```

## 2. Build the AgentSeurat

```{r ingest}
SeuratData::InstallData("pbmc3k")
seu <- SeuratData::LoadData("pbmc3k")
seu$sample <- "pbmc3k"

obj <- AgentSeurat(
  seu,
  initial_script = 'seu <- SeuratData::LoadData("pbmc3k")'
)
obj
```

The `obj` printout reports the scAgentKit version, the stage
(`initialized`), data shape, and zero recorded decisions so far.

## 3. QC

```{r qc}
obj <- qc_add_metrics(obj, species = "human")
obj <- qc_plot(obj, tag = "before", group_by = "sample")

obj <- qc_split(obj, split_by = "sample")
obj <- qc_threshold(obj, min_nCount = 1000, min_nFeature = 500,
                    max_percent_mt = 20)

# MAD filter: log10 space on count/feature (v0.2.0 default), linear on
# percent.mt (upper-only)
obj <- qc_mad(obj, nmad = 3)

obj <- qc_doublet(obj, remove = TRUE, seed = 999)
obj <- qc_remove_genes(obj, species = "human")
obj <- qc_merge(obj, join_layers = TRUE)
obj <- qc_plot(obj, tag = "after", group_by = "sample")
```

## 4. Normalize, HVG, scale, PCA

```{r pp}
obj <- sc_normalize(obj)
obj <- sc_find_variable_features(obj, nfeatures = 2000)
obj <- sc_scale(obj, vars_to_regress = c("nCount_RNA", "percent.mt"))
obj <- sc_pca(obj, npcs = 50)
```

## 5. LLM-driven PC selection

```{r pcs}
chat_fn <- chat_claude()    # uses ANTHROPIC_API_KEY env var
# Or: chat_fn <- chat_deepseek(); chat_grok(); chat_openai(); etc.

obj <- sc_select_pcs_visual(
  obj,
  chat_fn = chat_fn,
  tissue  = "PBMC"
)
```

The LLM is shown elbow / cumulative variance plots at multiple
candidate `ndim` values, and picks the one with the best
biology-respects-statistics tradeoff. The reasoning is logged.

## 6. UMAP, neighbors, cluster sweep, resolution recommend

```{r cluster}
obj <- sc_neighbors(obj)
obj <- sc_umap(obj)
obj <- sc_cluster_sweep(obj, resolutions = c(0.3, 0.5, 0.8, 1.0, 1.2))

obj <- sc_resolution_recommend(
  obj,
  chat_fn = chat_fn,
  tissue  = "PBMC",
  vision  = TRUE
)

# Commit the recommended resolution as `seurat_clusters`
obj <- sc_cluster_commit(obj)
```

## 7. Markers and reference matching

```{r markers}
obj <- sc_markers(obj)
obj <- sc_markers_summary(obj, top_n = 30)
obj <- annot_query_cellmarker(obj, species = "human", tissue = "Peripheral blood")
obj <- annot_match_reference(obj)
```

## 8. LLM annotation with anti-hallucination + ensemble + hybrid confidence

```{r annot}
# Ensemble mode: 3 independent LLM calls per cluster, majority vote.
# Requires temperature > 0 in chat_fn; rebuild with temperature = 0.4
# for diversity.
chat_fn_ens <- chat_claude(temperature = 0.4)

# Optional parallelism — set future::plan() first
# future::plan(future::multisession, workers = 4)

obj <- annot_llm_annotate(
  obj,
  chat_fn          = chat_fn_ens,
  tissue           = "PBMC",
  n_samples        = 3,
  validate_markers = TRUE,
  parallel         = FALSE       # set TRUE after future::plan()
)

obj <- annot_clean_celltypes(obj)
obj <- annot_apply(obj)
```

Each row of `obj@params$llm_annotations` now has:

- `primary_annotation`, `confidence` (LLM self-report)
- `hybrid_confidence`, `hybrid_confidence_label` (objective)
- `confidence_disagreement` (TRUE when LLM and hybrid land in opposite bands)
- `supporting_markers`, `contradicting_markers`
- `hallucinated_markers` (cited genes not in the marker evidence)
- `ensemble_agreement`, `ensemble_n` (modal-label fraction)
- `recommended_action` (`accept` / `flag_for_review` / `reject` / `mark_unknown`)

## 9. Optional: subcluster the broad types

```{r subcluster}
obj <- annot_subcluster(
  obj,
  chat_fn               = chat_fn,
  target                = NULL,      # all broad types meeting threshold
  subcluster_resolution = "adaptive",
  n_hvg                 = "adaptive",
  n_pcs                 = "adaptive",
  tissue                = "PBMC"
)
```

## 10. Optional: map annotations to Cell Ontology

```{r cl}
# install.packages("ontologyIndex")
obj <- annot_map_to_cl(obj, map_cells = TRUE)

# Interactive lookup if a label didn't match
cl_lookup("hepatocyte", n = 5)
```

## 11. Inspect the audit trail

```{r inspect}
# Decision log
get_decisions(obj)

# Full reproducible R script — re-running this from raw counts produces
# the same analysis (LLM-call non-determinism aside)
cat(get_script(obj))

# Token usage breakdown per pipeline step
get_token_usage(obj)

# Or the global running total
token_usage_summary()
```

## 12. Export

```{r export}
# HTML report with decisions, figures, annotations (flagged rows
# highlighted), token usage
report_html(obj, file = "pbmc3k_report.html")

# Persistable checkpoint
save_checkpoint(obj, "pbmc3k_final.qs")
```

## Where to go next

- For tissue-specific work, look at `inst/examples/full_pipeline_example.R`
  (mouse colon CRC + immunotherapy) and the HCC examples.
- The `benchmark/` directory contains scripts comparing scAgentKit's
  upstream decisions (PC selection, resolution recommendation, batch
  variable identification) against analyst-curated ground truth.
- See `?annot_llm_annotate` for the full set of ensembling and
  validation options.
